---
title: "STAT 183 Individual Dataset"
author: "Chuqi(Angel) Jin"
date: "2024-05-08"
output:
  pdf_document: 
    toc: yes
  html_document:
    df_print: paged
---
***
```{r,message=FALSE,warning=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(car)
library(mltools)
library(data.table)
library(fastDummies)
library(lmtest)
library(tidyr)
library(grid)
```

## Loading
```{r}
realEstate <- read.csv("D:/angelClass/ARclass/STAT_183/Individual/RealEstatePrice.csv") 
```

```{r}
# Subset again to get the useable dataset for the model building
# Replace all the empty value and 0 values into NA
realEstate[realEstate == ""] = NA
realEstate[realEstate == "?"] = NA
realEstate[realEstate == 0] = NA

# Clean the data set with selected restrictions
cleaned <- realEstate %>%
  na.omit() %>%
  filter(Year >= 2017) 

cleaned$Estimated.Value <- as.numeric(as.character(cleaned$Estimated.Value)) / 1000
cleaned$Sale.Price <- as.numeric(as.character(cleaned$Sale.Price)) / 1000
```

2574 observation

# DATA DESCIRPTION
# EDA 
```{r}
summary(cleaned)
sd(cleaned$Year)
```

## Date (Not include in the model building)
```{r}
# Identify which month sells most
date_eda <- data.frame(date = as.Date(c(cleaned$Date)))
date_eda$month <- format(date_eda$date, "%m")
p1 <- ggplot(date_eda, aes(x=month)) + 
  geom_bar(fill = "#7d9ea2", color = "white") +
  labs(x = "Months", y = "count")

# Identify which date sells most
date_eda$day <- format(date_eda$date, "%d")
p2 <- ggplot(date_eda, aes(x=day)) + 
  geom_bar(fill = "#d99b77", color = "white") +
  labs(x = "Days", y = "count")

grid.arrange(p1,p2, ncol = 1)
```

## Year (Include in the model building, numerical)
```{r}
# Identify which year sells most
p3 <- ggplot(cleaned, aes(x=Year)) + 
  geom_histogram(fill = "#7d9ea2", color = "white", binwidth=1) +
  labs(x = "Years", y = "count")
p3

y_freq <- table(cleaned$Year)
y_rank <- sort(y_freq, decreasing = T)
y_rank

year = c(2017, 2018, 2019, 2020, 2021, 2022)
count = c(426, 325, 488, 408, 484, 443)

year_set = data.frame(year, count)

ggplot(year_set, aes(x=year, y=count)) +
  geom_line(color="black")+
  geom_point()
```


## Locality (include in the model building, categorical)
```{r}
# Identify which year sells most
p4 <- ggplot(cleaned, aes(x=Locality, fill=Locality)) + 
  geom_bar(width=1) +
  labs(x = "Localities") +
  coord_polar()+
  theme(panel.background = element_blank())
p4
```
Bridgeport and Waterbury are top 2 location of property transaction

## Estimated Value(include in the model building, numerical)
```{r}
p5 <-ggplot(cleaned, aes(x = Estimated.Value)) + 
  geom_histogram(fill = "#7d9ea2", color="white")+
  labs(x = "Estimated Value (in thousand dollars)") 
p5
```

## Sale Price(include in the model building, Response variable)
```{r}
p6 <-ggplot(cleaned, aes(x = Sale.Price)) + 
  geom_histogram(fill = "#f26835", color="white")+
  labs(x = "Sale Price (in thousand dollars)", y = "count") 
p6
```
Not normally distributed, it's right skewed.


## Property(include in the model building, categorical)
```{r}
p7 <- ggplot(cleaned, aes(x = Property)) +
  geom_bar(fill = "#7d9ea2", color="white")

```
Single family and two family are top 2 property types

## Residential (include in the model building, categorical)
```{r}
p8 <- ggplot(cleaned, aes(x = Residential)) +
  geom_bar(fill = "#7d9ea2", color="white")
  
grid.arrange(p7,p8, ncol = 1)
```
Detached House and Duplex are the top 2

## Number of rooms (Include, numerical)
```{r}
p9 <- ggplot(cleaned, aes(x = num_rooms)) +
  geom_histogram(fill = "#7d9ea2", color="white",binwidth = 1)+
  labs(x = "Number of Rooms", y = "count")
  
```


## Number of bathrooms (Include, numerical)
```{r}
p10 <- ggplot(cleaned, aes(x = num_bathrooms)) +
  geom_histogram(fill = "#d99b77", color="white",binwidth = 1)+
  labs(x = "Number of Bathrooms", y = "count")
grid.arrange(p9,p10, ncol = 1)
```


## carpet area (Include, numerical)
```{r}
p11 <- ggplot(cleaned, aes(x = carpet_area)) + 
  geom_histogram(fill = "#d99b77", color="white", binwidth=300)+
  labs(x = "Carpet Area (in square feet)", y = "count")
p11
```

## Property tax rate (Include, numerical)
```{r}
p12 <- ggplot(cleaned, aes(x = property_tax_rate)) + 
  geom_histogram(fill = "#d99b77", color="white", binwidth=0.25)+
  labs(x = "Property Tax Rate", y = "count")
p12
```

## Face (Include, categorical)
```{r}
p13 <-ggplot(cleaned, aes(x = Face)) + 
  geom_bar(fill = "#d99b77", color="white")
p13
```

```{r}
# Face
# Find its frequency of Top 2 face direction
face_freq <- table(cleaned$Face)
face_rank <- sort(face_freq, decreasing = T)
face_rank
```
North and West are top 2 facing directions


```{r}
summary(cleaned)
sd(cleaned$Year)
sd(cleaned$Estimated.Value)
sd(cleaned$Sale.Price)
```



# Model Building Data
```{r}
modelData <- subset(cleaned, select = c("Year", "Locality", "Estimated.Value",
                                      "Property", "Residential", 
                                      "num_rooms", "num_bathrooms",
                                      "carpet_area", "property_tax_rate", 
                                      "Face", "Sale.Price"))


# Subset useful data for model building
modelData <- modelData %>%
  filter(Locality == "Bridgeport" | Locality == "Waterbury")%>%
  filter(Property == "Single Family" | Property == "Two Family") %>%
  filter(Residential == "Detached House" | Residential == "Duplex")%>%
  filter(Face == "North" | Face == "West") 

```

The subset dataset contains 397 observations, 11 variables
 
## Response vs Predictors
```{r, message=FALSE, warning=FALSE}
# create graphs between each predictor var and the response var
s.year <- ggplot(modelData, aes(x = Year, y = Sale.Price)) +
  geom_point(color = "#990066") 
s.estivalue <- ggplot(modelData, aes(x = Estimated.Value, y = Sale.Price)) +
  geom_point(color = "#990066") 
s.numroom <- ggplot(modelData, aes(x = num_rooms, y = Sale.Price)) +
  geom_point(color = "#990066") 
s.numbath <- ggplot(modelData, aes(x = num_bathrooms, y = Sale.Price)) +
  geom_point(color = "#990066") 
s.caparea <- ggplot(modelData, aes(x = carpet_area, y = Sale.Price)) +
  geom_point(color = "#990066") 
s.proprate <- ggplot(modelData, aes(x = property_tax_rate, y = Sale.Price)) +
  geom_point(color = "#990066") 

box.property <- ggplot(modelData, aes(x = Property, y = Sale.Price)) +
           geom_boxplot(color = "#990066", outlier.color = "red",
                        outlier.shape = 4, outlier.size = 2) 
box.residential <- ggplot(modelData, aes(x = Residential, y = Sale.Price)) +
           geom_boxplot(color = "#990066", outlier.color = "red",
                        outlier.shape = 4, outlier.size = 2) 
box.Locality <- ggplot(modelData, aes(x = Locality, y = Sale.Price)) +
           geom_boxplot(color = "#990066", outlier.color = "red",
                        outlier.shape = 4, outlier.size = 2)
box.face <- ggplot(modelData, aes(x = Face, y = Sale.Price)) +
           geom_boxplot(color = "#990066", outlier.color = "red",
                        outlier.shape = 4, outlier.size = 2)

# divide the frame in grid using grid.arrange() and put above plots in it
grid.arrange(s.estivalue, s.numroom, s.numbath, s.caparea, s.proprate,
             s.year, box.property, box.property, box.residential, box.Locality,
             box.face, ncol = 3)
```

## Multicollinearity Analysis
```{r fig.cap="Pairwise Comparison of the Original Data", message=FALSE, warning=FALSE}
ggpairs(modelData) + theme(axis.text.x = element_text(size = 5, angle = 90),
                           axis.text.y = element_text(size = 5),
                           strip.text.y = element_text(angle = 0))
```

```{r}
# All predictors
property.mlr <- lm(Sale.Price ~ Year+Locality+Estimated.Value+num_rooms+ Property+num_bathrooms+carpet_area+property_tax_rate+Face+Residential, 
data = modelData)

summary(property.mlr)

# Can not do the vif
# vif(property.mlr)
```


```{r}
# Remove property and residential those 2 variables
property1.mlr <- lm(Sale.Price ~ Year+Locality+Estimated.Value+num_rooms+num_bathrooms+carpet_area+property_tax_rate+Face, 
data = modelData)

summary(property1.mlr)

```

```{r}
vif(property1.mlr)
```

```{r}
# remove 'num_rooms' predictor
property2.mlr <- lm(Sale.Price ~ Year+Locality+Estimated.Value+num_bathrooms+carpet_area+property_tax_rate+Face, 
data = modelData)

summary(property2.mlr)

```

```{r}
vif(property2.mlr)
```
No more multicollinearity, all less than 5

### Stop for EDA

```{r}
# Model building dataset
modelbuilding <- subset(modelData, select = c("Year", "Locality", "Estimated.Value"
                                        , "num_bathrooms", "carpet_area", "property_tax_rate", 
                                      "Face", "Sale.Price"))

```

397 obs with 8 variables

### Forward Stepwise Regression 

```{r, message=FALSE, warning=FALSE}
# perform stepwise regression using the original data set

# specify the null model with no predictor
original.null <- lm(Sale.Price ~ 1, data = modelbuilding)
# specify the full model using all of the potential predictors
original.full <- lm(Sale.Price ~ ., data = modelbuilding)
# use the stepwise algorithm to build a parsimonious model
original.step <- step(original.null,
                      scope = list(lower = original.null, upper = original.full),
                      direction = "both", test = "F")

summary(original.step)
```

# Residual Diagnostics
```{r}
# check linearity
ggplot(data = original.step, aes(x = .fitted, y = .resid)) +
  geom_point(alpha = 0.5, color = "darkblue") +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(x = "Predicted Values", y = "Residuals")
# There is a pattern, so we look at residuals vs. each predictors

# predictor: Estimated.Value
l.ev <- ggplot(data = original.step, aes(x = Estimated.Value, y = .resid)) +
  geom_point(alpha = 0.5, color = "darkblue") +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(x = "Estimated.Value", y = "Residuals")
# predictor: Year
l.year <- ggplot(data = original.step, aes(x = Year, y = .resid)) +
  geom_point(alpha = 0.5, color = "darkblue") +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(x = "Years", y = "Residuals")
# predictor: Locality
l.Local <- ggplot(data = original.step, aes(x = Locality, y = .resid)) +
  geom_point(alpha = 0.5, color = "darkblue") +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(x = "Locality", y = "Residuals")
# predictor: Carpet Area
l.cap <- ggplot(data = original.step, aes(x = carpet_area, y = .resid)) +
  geom_point(alpha = 0.5, color = "darkblue") +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(x = "Carpet Area", y = "Residuals")

# divide the frame in grid using grid.arrange() and put above plots in it
grid.arrange(l.ev, l.year, l.Local, l.cap, ncol = 2, nrow = 2)
```

```{r fig.cap="Residual Diagnostics: Constant Variance", message=FALSE, warning=FALSE}
# perform residual diagnostics on the residuals of the final stepwise model

# check constant variance
bptest(original.step)
```

p-value is extremly small, therefore, there is
sufficient evidence to conclude the residuals do not have constant variance.

```{r fig.cap="Residual Diagnostics: Normality", message=FALSE, warning=FALSE}
# perform residual diagnostics on the residuals of the final stepwise model

# check normality 
ggplot(data = original.step, aes(x = .resid)) + 
  geom_histogram(fill = "darkblue", color = "white") +
  labs(x = "revenue (in thousand $)", y = "count")

ggplot(data = original.step, aes(x = .resid)) + 
  geom_boxplot(color = "darkblue", outlier.color = "red",
               outlier.shape = 4, outlier.size = 2) +
           labs(y = "revenue (in thousand $)")

qqnorm(resid(original.step))
qqline(resid(original.step))

shapiro.test(resid(original.step))
```
Even though out histogram looks bell-shaped, but shapiro-wilk test p-value is too small, so normlaity also not satisfied

```{r, message=FALSE, warning=FALSE}
# perform residual diagnostics on the residuals of the final stepwise model

# check independence
durbinWatsonTest(original.step)
```
p-value is > 0.05, therefore, independence assumption is satisfied

## Model Building and Diagnostics

```{r, message=FALSE, warning=FALSE}

sp <- modelbuilding$Sale.Price
year <- modelbuilding$Year
ev <- modelbuilding$Estimated.Value
nb <- modelbuilding$num_bathrooms
ca <- modelbuilding$carpet_area
ptr <- modelbuilding$property_tax_rate

# create dummy variables for each cat var
L <- ifelse(modelbuilding$Locality == "Waterbury", 1, 0)
face <- ifelse(modelbuilding$Face == "West", 1, 0)
```

```{r fig.cap="Initial Plots of Predictor vs. Response Variables", message=FALSE, warning=FALSE}
# check the relationship between each predictor var and rev
a <- ggplot(modelbuilding, aes(x = Year, y = sp)) +
  geom_point(color = "darkblue") 

b <- ggplot(modelbuilding, aes(x = Estimated.Value, y = sp)) +
  geom_point(color = "darkblue") 
 
c <- ggplot(modelbuilding, aes(x = num_bathrooms, y = sp)) +
  geom_point(color = "darkblue") 

d <- ggplot(modelbuilding, aes(x = carpet_area, y = sp)) +
  geom_point(color = "darkblue") 

e <- ggplot(modelbuilding, aes(x = property_tax_rate, y = sp)) +
  geom_point(color = "darkblue") 

f <- ggplot(modelbuilding, aes(x = Locality, y = sp)) +
  geom_point(color = "darkblue") 

g <- ggplot(modelbuilding, aes(x = Face, y = sp)) +
  geom_point(color = "darkblue") 

# divide the frame in grid using grid.arrange() and put above plots in it
grid.arrange(a, b, c, d, e, f, g, ncol = 3, nrow = 3)
```

### Model Building

```{r fig.cap="Interaction Between popularity and genres", message=FALSE, warning=FALSE}
# check interaction between 
plot(ev, sp, type = "n")
points(ev[L == 0], sp[L == 0], col="red")
points(ev[L == 1], sp[L == 1], col="green")
```

interaction yes!!!!!!

```{r fig.cap="Residual Diagnostics: y1", message=FALSE, warning=FALSE}
lm1 <- lm(sp ~ 0 + ev * L)
y1 <- lm1$residuals

ggplot(modelbuilding, aes(x = y1))+
  geom_histogram(fill = "darkblue", color = "white") +
  labs(x = "revenue (in thousand $)", y = "count")
qqnorm(y1)
qqline(y1)
shapiro.test(y1)
```
Normality failed

```{r fig.cap="Plots of Predictors vs. y1", message=FALSE, warning=FALSE}
# check the relationship between each predictor variable and y1
a <- ggplot(modelbuilding, aes(x = Year, y = y1)) +
  geom_point(color = "darkblue") 

b <- ggplot(modelbuilding, aes(x = Estimated.Value, y = y1)) +
  geom_point(color = "darkblue") 
 
c <- ggplot(modelbuilding, aes(x = num_bathrooms, y = y1)) +
  geom_point(color = "darkblue") 

d <- ggplot(modelbuilding, aes(x = carpet_area, y = y1)) +
  geom_point(color = "darkblue") 

e <- ggplot(modelbuilding, aes(x = property_tax_rate, y = y1)) +
  geom_point(color = "darkblue") 

f <- ggplot(modelbuilding, aes(x = Locality, y = y1)) +
  geom_point(color = "darkblue") 

g <- ggplot(modelbuilding, aes(x = Face, y = y1)) +
  geom_point(color = "darkblue") 

# divide the frame in grid using grid.arrange() and put above plots in it
grid.arrange(a, b, c, d, e, f, g, ncol = 3, nrow = 3)
```


carpet area, num_bathrooms seems no pattern

```{r fig.cap="Residual Diagnostics: y2", message=FALSE, warning=FALSE}
lm2 <- lm(y1 ~ 0 + year)
y2 <- lm2$residuals

ggplot(modelbuilding, aes(x = y2))+
  geom_histogram(fill = "darkblue", color = "white") 
  
qqnorm(y2)
qqline(y2)
shapiro.test(y2)
```


```{r fig.cap="Plots of Predictors vs. y1", message=FALSE, warning=FALSE}
# check the relationship between each predictor variable and y1
a <- ggplot(modelbuilding, aes(x = Year, y = y2)) +
  geom_point(color = "darkblue") 

b <- ggplot(modelbuilding, aes(x = Estimated.Value, y = y2)) +
  geom_point(color = "darkblue") 
 
c <- ggplot(modelbuilding, aes(x = num_bathrooms, y = y2)) +
  geom_point(color = "darkblue") 

d <- ggplot(modelbuilding, aes(x = carpet_area, y = y2)) +
  geom_point(color = "darkblue") 

e <- ggplot(modelbuilding, aes(x = property_tax_rate, y = y2)) +
  geom_point(color = "darkblue") 

f <- ggplot(modelbuilding, aes(x = Locality, y = y2)) +
  geom_point(color = "darkblue") 

g <- ggplot(modelbuilding, aes(x = Face, y = y2)) +
  geom_point(color = "darkblue") 

# divide the frame in grid using grid.arrange() and put above plots in it
grid.arrange(a, b, c, d, e, f, g, ncol = 3, nrow = 3)
```

Not much changes, log transformation of response

```{r fig.cap="Plots of Predictor vs. lnrev", message=FALSE, warning=FALSE}
# perform log transformation on revenue and edit the original data set to
# exclude revenue and include lnrev
lnsp <- log(modelbuilding$Sale.Price)
modelbuilding$lnsp <- lnsp
modelbuilding <- modelbuilding[, -8]

# convert cat vars genres and original language to be binary data
modelbuilding$Locality <- L
modelbuilding$Face <- face

# check the relationship between each predictor variable and lnrev
a <- ggplot(modelbuilding, aes(x = Year, y = lnsp)) +
  geom_point(color = "darkblue") 

b <- ggplot(modelbuilding, aes(x = Estimated.Value, y = lnsp)) +
  geom_point(color = "darkblue") 
 
c <- ggplot(modelbuilding, aes(x = num_bathrooms, y = lnsp)) +
  geom_point(color = "darkblue") 

d <- ggplot(modelbuilding, aes(x = carpet_area, y = lnsp)) +
  geom_point(color = "darkblue") 

e <- ggplot(modelbuilding, aes(x = property_tax_rate, y = lnsp)) +
  geom_point(color = "darkblue") 

f <- ggplot(modelbuilding, aes(x = Locality, y = lnsp)) +
  geom_point(color = "darkblue") 

g <- ggplot(modelbuilding, aes(x = Face, y = lnsp)) +
  geom_point(color = "darkblue") 

# divide the frame in grid using grid.arrange() and put above plots in it
grid.arrange(a, b, c, d, e, f, g, ncol = 3, nrow = 3)
```

recheck interaction
```{r fig.cap="Interaction Between popularity and genres", message=FALSE, warning=FALSE}
# check interaction between 
plot(ev, lnsp, type = "n")
points(ev[L == 0], lnsp[L == 0], col="darkred")
points(ev[L == 1], lnsp[L == 1], col="darkblue")
```

Similar to the results from before, the two variables overlap, suggesting a
strong interaction between them.

```{r fig.cap="Residual Diagnostics: y1", message=FALSE, warning=FALSE}
lm1 <- lm(lnsp ~ 0 + ev * L)
y1 <- lm1$residuals

ggplot(modelbuilding, aes(x = y1))+
  geom_histogram(fill = "darkblue", color = "white") +
  labs(x = "lnsp (in thousand $)", y = "count")
qqnorm(y1)
qqline(y1)
shapiro.test(y1)
```

No normality 

```{r}
# check the relationship between each predictor variable and lnrev
a <- ggplot(modelbuilding, aes(x = Year, y = y1)) +
  geom_point(color = "darkblue") 

b <- ggplot(modelbuilding, aes(x = Estimated.Value, y = y1)) +
  geom_point(color = "darkblue") 
 
c <- ggplot(modelbuilding, aes(x = num_bathrooms, y = y1)) +
  geom_point(color = "darkblue") 

d <- ggplot(modelbuilding, aes(x = carpet_area, y = y1)) +
  geom_point(color = "darkblue") 

e <- ggplot(modelbuilding, aes(x = property_tax_rate, y = y1)) +
  geom_point(color = "darkblue") 

f <- ggplot(modelbuilding, aes(x = Locality, y = y1)) +
  geom_point(color = "darkblue") 

g <- ggplot(modelbuilding, aes(x = Face, y = y1)) +
  geom_point(color = "darkblue") 

# divide the frame in grid using grid.arrange() and put above plots in it
grid.arrange(a, b, c, d, e, f, g, ncol = 3, nrow = 3)
```

everthing is good except carpet_area --> no patern


```{r fig.cap="Residual Diagnostics: y2", message=FALSE, warning=FALSE}
lm2 <- lm(y1 ~ 0 + year)
y2 <- lm2$residuals

ggplot(modelbuilding, aes(x = y2))+
  geom_histogram(fill = "darkblue", color = "white") 
  
qqnorm(y2)
qqline(y2)
shapiro.test(y2)
```

```{r}
# check the relationship between each predictor variable and lnrev
a <- ggplot(modelbuilding, aes(x = Year, y = y2)) +
  geom_point(color = "darkblue") 

b <- ggplot(modelbuilding, aes(x = Estimated.Value, y = y2)) +
  geom_point(color = "darkblue") 
 
c <- ggplot(modelbuilding, aes(x = num_bathrooms, y = y2)) +
  geom_point(color = "darkblue") 

d <- ggplot(modelbuilding, aes(x = carpet_area, y = y2)) +
  geom_point(color = "darkblue") 

e <- ggplot(modelbuilding, aes(x = property_tax_rate, y = y2)) +
  geom_point(color = "darkblue") 

f <- ggplot(modelbuilding, aes(x = Locality, y = y2)) +
  geom_point(color = "darkblue") 

g <- ggplot(modelbuilding, aes(x = Face, y = y2)) +
  geom_point(color = "darkblue") 

# divide the frame in grid using grid.arrange() and put above plots in it
grid.arrange(a, b, c, d, e, f, g, ncol = 3, nrow = 3)
```

```{r fig.cap="Residual Diagnostics: y3", message=FALSE, warning=FALSE}
lm3 <- lm(y2 ~ 0 + nb)
y3 <- lm3$residuals

ggplot(modelbuilding, aes(x = y3))+
  geom_histogram(fill = "darkblue", color = "white") +
  labs(x = "lnrev (in thousand $)", y = "count")
qqnorm(y3)
qqline(y3)
shapiro.test(y3)
```

```{r}
# check the relationship between each predictor variable and lnrev
a <- ggplot(modelbuilding, aes(x = Year, y = y3)) +
  geom_point(color = "darkblue") 

b <- ggplot(modelbuilding, aes(x = Estimated.Value, y = y3)) +
  geom_point(color = "darkblue") 
 
c <- ggplot(modelbuilding, aes(x = num_bathrooms, y = y3)) +
  geom_point(color = "darkblue") 

d <- ggplot(modelbuilding, aes(x = carpet_area, y = y3)) +
  geom_point(color = "darkblue") 

e <- ggplot(modelbuilding, aes(x = property_tax_rate, y = y3)) +
  geom_point(color = "darkblue") 

f <- ggplot(modelbuilding, aes(x = Locality, y = y3)) +
  geom_point(color = "darkblue") 

g <- ggplot(modelbuilding, aes(x = Face, y = y3)) +
  geom_point(color = "darkblue") 

# divide the frame in grid using grid.arrange() and put above plots in it
grid.arrange(a, b, c, d, e, f, g, ncol = 3, nrow = 3)
```

Take out carpet area, nb and face

```{r fig.cap="Residual Diagnostics: lm5", message=FALSE, warning=FALSE}
# fit first full model
lm4 <- lm(lnsp ~ 0+year+ev*L+ptr)
summary(lm4)

# check the residual assumptions
ggplot(modelbuilding, aes(x = lm4$residuals))+
  geom_histogram(fill = "darkblue", color = "white") +
  labs(x = "lnrev (in thousand $)", y = "count")
qqnorm(lm4$residuals)
qqline(lm4$residuals)
ggplot(modelbuilding, aes(x = lm4$fitted.values, y = lm4$residuals)) +
  geom_point(color = "darkblue") +
  labs(x = "Fitted Values", y = "Residuals")

shapiro.test(lm4$residuals)
durbinWatsonTest(lm4)
bptest(lm4)
```


```{r}
m2 <- lm(lnsp ~ 0+year+ev*L)
summary(m2)
```


```{r fig.cap="Pairwise Comparison for Model 4", message=FALSE, warning=FALSE}
# edit the data set to exclude runtime and vote average
modelbuilding <- modelbuilding[, -c(4, 5, 7)]

# create pairwise scatter plots and coefficient of correlation
ggpairs(modelbuilding) + theme(axis.text.x = element_text(size= 2.5, angle = 90),
                      axis.text.y = element_text(size = 2.5),
                      strip.text.y = element_text(angle = 0))

```

We still have multicollinearity issues in the dataset
```{r, message=FALSE, warning=FALSE}
# perform forward stepwise regression and use this model as Model 6

# specify a null model with no predictors
null.model <- lm(lnsp ~ 0, data = modelbuilding)

# use a stepwise algorithm to build a parsimonious model
# null to full
lm5 <- step(null.model,
            scope = list(lower = null.model, upper = lm4),
            direction = "both", test = "F")
summary(lm5)

# check the residual assumptions
ggplot(modelbuilding, aes(x = lm5$residuals))+
  geom_histogram(fill = "darkblue", color = "white") +
  labs(x = "lnrev (in thousand $)", y = "count")
qqnorm(lm5$residuals)
qqline(lm5$residuals)
ggplot(modelbuilding, aes(x = lm5$fitted.values, y = lm5$residuals)) +
  geom_point(color = "darkblue") +
  labs(x = "Fitted Values", y = "Residuals")

shapiro.test(lm5$residuals)
durbinWatsonTest(lm5)
bptest(lm5)
```
Same results, so we choose model 5 as final result

```{r}
# Model Comparison

# Model 1: All predictors in linear form
m1 <- lm(lnsp ~ 0+year+ev+L+ptr)
summary(m1)

# Model 2: 
m2 <- lm(lnsp ~ 0+year+ev*L+ptr)
summary(m2)

```

```{r}
anova(m1, m2)
```

```{r}
vif(m2)
```



```{r}
# check the residual assumptions
ggplot(modelbuilding, aes(x = m1$residuals))+
  geom_histogram(fill = "darkblue", color = "white") +
  labs(x = "lnrev (in thousand $)", y = "count")
qqnorm(m1$residuals)
qqline(m1$residuals)
ggplot(modelbuilding, aes(x = m1$fitted.values, y = m1$residuals)) +
  geom_point(color = "darkblue") +
  labs(x = "Fitted Values", y = "Residuals")

shapiro.test(m1$residuals)
durbinWatsonTest(m1)
bptest(m1)
```


### Final Model Diagnostics

```{r, message=FALSE, warning=FALSE}
final.model <- lm(lnsp ~ 0 + ptr + ev * L + year)

# non-significant F-tests with overall significant F-test
summary(final.model)

# variance inflation factor (VIF)
vif(final.model)

# check the residual assumptions
ggplot(modelbuilding, aes(x = final.model$residuals))+
  geom_histogram(fill = "darkblue", color = "white") +
  labs(x = "lnrev (in thousand $)", y = "count")

qqnorm(final.model$residuals)
qqline(final.model$residuals)
ggplot(modelbuilding, aes(x = final.model$fitted.values, y = final.model$residuals)) +
  geom_point(color = "darkblue") +
  labs(x = "Fitted Values", y = "Residuals")

shapiro.test(final.model$residuals)
durbinWatsonTest(final.model)
bptest(final.model)
```
